
---
title: "Field edge management practice survey data"
date: "October 26, 2015"
output: pdf_document
---

**0. MISSING DATA: MULTIPLE IMPUTATION**


```{r}
setwd("e:/Lectures3/488-CL/client")
ca <- read.csv("data1110.csv")
library(mice)

#data with all variables included in model
newca <- ca[c(9:11,24:37,39:51,53,55,57,59,61,66,68,70,90,91,93,96,98:102,105)]

#dataset with main predictors, ca_categorical for categorical/binary variables
#ca_ordinal for ordinal variables, ca_cont for continuous variables
ca_categorical <- ca[c(9:11,24:37,39:51,53,55,57,59,61,91,93,96,98:102,105)]
ca_ordinal <- ca[c(66,68,70)]
ca_cont <- ca[c(9,90)] #ignore variable hedgerow here, this is to avoid error message due to one variables included only

#factor all categorical/binary variables and order the ordinal variables
#categorical/binary variables factor
var_names_cat <- c("hedgerow","wildflower","perennial","X3a","X3b","X3c","X3d", 
"X3e","X3f","X3g","X3h","X3i","X3j","X3k","X3l","X3m","X3n","X3p","X3q","X3r",
"X3s","X3t","X3u","X3v","X3w","X3x","X3y","X3z","X3aa","X3ab","UCCE","RCD","NGO","CC",
"Agencies","A3","A4","A6","A7_owner","A7_technical","A7_bachelor","A7_advanced","A8","A9")
for (i in 1:44){
  ca_categorical[,var_names_cat[i]] <- factor(ca_categorical[,var_names_cat[i]])
}

#factor all ordinal variables
var_names_ord <- c("X5q","X6q","Q7")
for (i in 1:3){
  ca_ordinal[,var_names_ord[i]] <- ordered(ca_ordinal[,var_names_ord[i]])
}


#imputing data
#categorical/binary: to fasten the process of imputation, we divide dataset into 11 variables
ca_categorical1 <- ca_categorical[c(1:11)]
ca_categorical2 <- ca_categorical[c(12:22)]
ca_categorical3 <- ca_categorical[c(23:33)]
ca_categorical4 <- ca_categorical[c(34:44)]

#imput cat/binary variables
imput_ca_categorical1 <- mice(ca_categorical1, me=c(rep("polyreg",11)))
imput_ca_categorical2 <- mice(ca_categorical2, me=c(rep("polyreg",11)))
imput_ca_categorical3 <- mice(ca_categorical3, me=c(rep("polyreg",11)))
imput_ca_categorical4 <- mice(ca_categorical4, me=c(rep("polyreg",5),rep("logreg",4),rep("polyreg",2)))


#ordinal
#lets do imputation without dividing the data
imput_ca_ordinal <- mice(ca_ordinal,me=c(rep("polr",3)))

#continuous
#lets just do imputation
ca_cont$hedgerow <- factor(ca_cont$hedgerow)
imput_ca_cont <- mice(ca_cont,me=c("polyreg","pmm"))


#complete data check
com_ca_categorical1 <- complete(imput_ca_categorical1,1)
com_ca_categorical2 <- complete(imput_ca_categorical2,1)
com_ca_categorical3 <- complete(imput_ca_categorical3,1)
com_ca_categorical4 <- complete(imput_ca_categorical4,1)
com_ca_ordinal <- complete(imput_ca_ordinal,1)
com_ca_cont <- complete(imput_ca_cont,1)


#combine all complete data
combine_all <- cbind(com_ca_categorical1,com_ca_categorical2,com_ca_categorical3,com_ca_categorical4,com_ca_ordinal,com_ca_cont)
#delete one variable hedgerow 
ca_complete <- combine_all[c(-48)]

#export ca_complete to xlsx
library(xlsx)
write.xlsx(ca_complete, "e:/Lectures3/488-CL/client/ca_complete.xlsx")

#there it goes our final dataset is ca_complete
#NOTE: age group (A6) and income (A8) are confusing whether nominal or ordinal, the presence of ND messed it up

```


**1. DESCRIPTIVE STATISTICS**

*a. Table and Bar*

- Frequency and contingency tables

Table 1. Number (percentage) of farm manager currently, tried, and never used practices on farm edges.
```{r}

setwd("e:/Lectures3/488-CL/client")
library(xlsx)
ca_complete <-  read.xlsx("ca_complete.xlsx",sheetName = "Sheet1")

#############################################################
#table 1: best reflect opinion on field edge mngt#

dtfr <- data.frame(Practices=c(),Current=c(),Tried=c(),Never=c())
field_tab1 <- c("hedgerow","wildflower","perennial")
for (i in 1:3){
  sum_tab1 <- as.numeric(table(ca_complete[field_tab1[i]]))
  relative_tab1 <- (as.numeric((table(ca_complete[field_tab1[i]]))/sum(!is.na(ca_complete[field_tab1[i]]))))
  rel_deci <- sprintf("%.2f",relative_tab1)
  d <- data.frame(Practice=field_tab1[i],Current=paste(sum_tab1[1],
  " (", rel_deci[1],")",sep=""),Tried=paste(sum_tab1[2],
  " (", rel_deci[2],")", sep=""),Never=paste(sum_tab1[3],
  " (",rel_deci[3],")",sep=""))
  dtfr <- rbind(dtfr,d)
}
print(dtfr)

#te <- sum(!is.na(ca[field_tab1[1]]))

#opinion_tab1 <- rbind(as.numeric(table(ca$mow)), as.numeric(table(ca$burn)))
#tab1 <- cbind(field_tab1,opinion_tab1)
```

Table 2. Number (percentage) of farm manager that currently, previously, never, and considering collaboration with partner/collabolator
```{r}
#############################################################
#table2
#print table agencies 
dtfr3 <- data.frame(Partner=c(),NeverUsed=c(),Considering=c(),Previously=c(),Currently=c())
field_tab3 <- c("UCCE","Agencies","RCD","CC","NGO")
for (i in 1:5){
  sum_tab3 <- as.numeric(table(factor(ca_complete[[field_tab3[i] ]],level=0:3)))
  relative_tab3 <- (as.numeric((table(factor
  (ca_complete[[field_tab3[i] ]],levels=0:3))))/
  sum(!is.na(ca_complete[field_tab3[i]])))
  rel_deci3 <- sprintf("%.2f",relative_tab3)
  d3 <- data.frame(Partner=field_tab3[i],
  NeverUsed=paste(sum_tab3[1]," (",rel_deci3[1],
  ")",sep=""),Considering=paste(sum_tab3[2],
  " (",rel_deci3[2],")",sep=""),Previously=paste(sum_tab3[3],
  " (" ,rel_deci3[3],")",sep=""),Currently=paste(sum_tab3[4],
  " (",rel_deci3[4],")",sep=""))
  dtfr3 <- rbind(dtfr3,d3)
}
print(dtfr3)
```

Table 3. Number of farm (percentage) based on categories of farm size (in acres) 

```{r}
#############################################################
#table3
#size of farm vs interest
#classify the farm size into some groups based 
#on categories from USDA National Agricultural
#Statistics Service (NASS)
breaks <- c(1,9,19,29,49,99,199,499,999,1999,1280000)
A2_cut <- cut(ca_complete$A2,breaks,right=FALSE)
number.farm <- table(A2_cut)
#relative freq
percentage <- table(A2_cut)/sum(!is.na(ca_complete$A2))
comma <- options(digits=1)
cbind(number.farm,percentage)

#Note: relative freq of quantitative data: http://www.r-tutor.com/elementary-statistics/quantitative-data/
#relative-frequency-distribution-quantitative-data
#############################################################
```

Table 4. Level of interest in field edges practices based on education and skills. 
NI=Not Interested, SI=Somewhat interested, VI=Very interested
```{r}
#############################################################
#table4
#print table interest vs edu+skills 
dtfr4 <- data.frame(Education_Skill=c(),NI=c(),SI=c(),VI=c())
  sum_tab4 <- table(subset(ca$Q7, ca$A7_technical == 1 & ca$A7_bachelor == 0 & ca$A7_advanced == 0) )
  sum_tab5 <- table(subset(ca$Q7, ca$A7_technical == 0 & ca$A7_bachelor == 1 & ca$A7_advanced == 0) )
  sum_tab6 <- table(subset(ca$Q7, ca$A7_technical == 0 & ca$A7_bachelor == 0 & ca$A7_advanced == 1) )
  sum_tab7 <- table(subset(ca$Q7, ca$A7_technical == 1 & ca$A7_bachelor == 1 & ca$A7_advanced == 0) )
  sum_tab8 <- table(subset(ca$Q7, ca$A7_technical == 1 & ca$A7_bachelor == 0 & ca$A7_advanced == 1) )
  
  d4 <- data.frame(Education_Skill="Technical",NI=paste(sum_tab4[2]),
  SI=paste(sum_tab4[3]),VI=paste(sum_tab4[4]))
  d5 <- data.frame(Education_Skill="Bachelor",NI=paste(sum_tab5[2]),
  SI=paste(sum_tab5[3]),VI=paste(sum_tab5[4]))
  d6 <- data.frame(Education_Skill="Advanced",NI=paste(sum_tab6[2]),
  SI=paste(sum_tab6[3]),VI=paste(sum_tab6[4]))
  d7 <- data.frame(Education_Skill="Technical+Bachelor",NI=paste(sum_tab7[2]),
  SI=paste(sum_tab7[3]),VI=paste(sum_tab7[4]))
  d8 <- data.frame(Education_Skill="Technical+Advanced",NI=paste(sum_tab8[2]),
  SI=paste(sum_tab8[3]),VI=paste(sum_tab8[4]))
  
  dtfr4 <- rbind(dtfr4,d4,d5,d6,d7,d8)
print(dtfr4)

#############################################################
```

Table 5(a). Number of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on age groups
```{r}
#############################################################

#age vs interest

#before that, lets fill the missing value in A6_new &
#Q7 with NA, so this wont confuse with "" or " " category

ca$A6_new[ ca$A6_new == ""  ] = NA
ca$Q7[ ca$Q7 == ""  ] = NA

ca$Q7 <- factor(ca$Q7)
ca$A6_new <- factor(ca$A6_new)
levels(ca$Q7) 
agi <- table(ca$A6_new,ca$Q7)
rownames(agi) 
addmargins(agi) #addmargins to adding row and col sums
```

Table 5(b). Percentage of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on age groups (percentage is per column of level of interest)

```{r}
prop.table(agi,2) # percentages per column
#note: contingency tables manipulation: #http://cnx.org/contents/d846ad27-dc3e-4a2e-9d90-90dc48688fc7@14.2:27
#############################################################
```

Table 6a. Number farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on gender
```{r}
#fill the missing value in A9 with NA, so this wont confuse with "" or " " category
ca$A9[ ca$A9 == ""  ] = NA #after labeling NA do
#not forget to run factor(ca$) so this can eliminate the "" or " "

#gender vs interest
ca$A9 <- factor(ca$A9)
ca$Q7 <- factor(ca$Q7)

levels(ca$A9)  
gei <- table(ca$A9,ca$Q7)
rownames(gei) 
addmargins(gei)

#############################################################
```

Table 6b. Percentage of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on gender (percentage is per column of level of interest)
```{r}
prop.table(gei,2) # percentages per column
```

Table 7a. Number farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on income
```{r}
#fill the missing value in A8_new with NA, so this wont confuse with "" or " " category
ca$A8_new[ ca$A8_new == ""  ] = NA

#income vs interest
ca$A8_new <- factor(ca$A8_new)
levels(ca$A8_new) 
ini <- table(ca$A8_new,ca$Q7)
rownames(ini) 
addmargins(ini)
```

Table 7b. Percentage of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on income (percentage is per column)

```{r}
prop.table(ini,2) # percentages per column
```

Table 8a. Number farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on familiarity with program that offer financial and technical support to incorporate field edge plantings
```{r}
#fill the missing value in X6q with NA, so this wont confuse with "" or " " category
ca$X6q[ ca$X6q == ""  ] = NA

#familiarity vs interest 
ca$X6q <- factor(ca$X6q)

levels(ca$X6q)
fi <- table(ca$X6q,ca$Q7)
rownames(fi) 
addmargins(fi)
```

Table 8b. Percentage of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on familiarity with program that offer financial and technical support to incorporate field edge plantings (percentage per column)

```{r}
prop.table(fi,2) # percentages per column
```

Table 9a.Number of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on actual application of the practice

```{r}
#fill the missing value in Q7 with NA, so this wont confuse with "" or " " category
ca$Q7[ ca$Q7 == ""  ] = NA

ca$Q7 <- factor(ca$Q7)
levels(ca$Q7)
Q7hrow <- table(ca$hedgerow,ca$Q7)
rownames(Q7hrow) <- c("never","tried","currently")
addmargins(Q7hrow)
```

Table 9b.Percentage of farm manager that NI(Not interested), SI(Somewhat interested),
and VI(Very interested) to field edge practices based on actual application of the practice (percentage per column)

```{r}
prop.table(Q7hrow,2) # percentages per column
```

- Bar charts/Visualization

*b. Calculate missing data in every column*

- Calculate NA 

- Calculate null 

```{r}
#colSums(is.na(ca))
```

*c. Testing for contingency table (chi-square, fisher)*

Conduct chi-square test and Fisher test (without missing data/null/NA) at 0.05 significance level.

- Chi square test below yield warning message related to expected value less than 5:


```{r}
library(MASS)  
chisq.test(agi) #Test hypothesis whether interest levels is independent with age group (Table 5a)
chisq.test(ini) #Test hypothesis whether interest levels is independent with income group (Table 7a)
chisq.test(gei) #Test hypothesis whether interest levels is independent with gender (Table 6a)
chisq.test(Q7hrow) #Test hypothesis whether interest levels is independent with actual application (Table 9a)

#Note: warning from chi-square test, alternative solution: 
#1. http://ww2.coastal.edu/kingw/statistics/R-tutorials/independ.html
#2. downsizing groups of variable
```

- Chi square test below yield no warning message:

```{r}
chisq.test(fi) #Test hypothesis whether interest levels is 
#independent with familiarity with program and financial support (Table 8a)

```

**2. MODEL BUILDING**

Applying logistic regression to estimate the probability of people who plant hedgerows, wildflowers, and native perennial grass.

```{r}
####################################################################
#predictor variables

#non demographic predictor variables
#1. benefit,calculate number of more benefit or high benefit divided by the number of all variables associated with benefit in every rows. the calculation yield one value for every records/rows, we call this as new variable as benefit
comma <- options(digits=1) #2 digit 
benefit <- (rowSums(ca_complete[,5:18] ==3 | ca_complete[,5:18] ==2, na.rm=T))/ncol(ca_complete[,5:18])
#divide outcome, use (=1) and not use the practice (=0). 

#2.concern, calculate number of more concern or high concern divided by the number of all variables associated with concern in every rows. the calculation yield one value for every records/rows, we call this as new variable as concern
comma <- options(digits=1) #2 digit 
concern <- (rowSums(ca_complete[,19:31] ==3 | ca_complete[,19:31] ==2, na.rm=T))/ncol(ca_complete[,19:31])

#3a.partnership, have partner(=1) are currently and previously, otherwise have no partner(=0)
new_UCCE <- as.numeric(ca_complete$UCCE==2 | ca_complete$UCCE==3)
new_RCD <- as.numeric(ca_complete$RCD==2 | ca_complete$RCD==3)
new_NGO <- as.numeric(ca_complete$NGO==2 | ca_complete$NGO==3)
new_CC <- as.numeric(ca_complete$CC==2 | ca_complete$CC==3)
new_Agencies <- as.numeric(ca_complete$Agencies==2 | ca_complete$Agencies==3)


#3b.partnership, have partner(=1) is currently, otherwise have no partner(=0)

#1. current use is classified as use (=1), tried and never use are classified as not use (=0).
use1_h <- as.numeric(ca_complete$hedgerow==2) #hedgerow
use1_f <- as.numeric(ca_complete$wildflower==2) #wildflower
use1_g <- as.numeric(ca_complete$perennial==2) #perennial

#2. current use and never use are classified as use (=1), and never use is classified as not use (=0).
use2_h <- as.numeric(!(ca$hedgerow==0))

#3. tried/discontinued classified as use in past (=1), and never use
#and currently used as not use in the past(=0) 
tried <- as.numeric(ca$hedgerow==1)

#create new table included use1_h and other predictors
caNew <- cbind(use1_h,new_UCCE,new_RCD,new_NGO,new_CC,new_Agencies,concern,benefit,ca_complete,use1_f,use1_g)
caNew_2 <- caNew[c(-9,-13:-44)]

#export caNew_2 to xlsx
library(xlsx)
write.xlsx(caNew_2, "e:/Lectures3/488-CL/client/caNew_2.xlsx")

####################################################################
```

a. Model 1 (current use as use(=1), others as not use(=0))

```{r}
####################################################################
#Variables in the model:

#Below is demographic variables:
#A2 -> farm size, continuous
#A3 -> ownership of farm, discrete char
#A4_new -> conventional/organic/both, discrete char
#A6_new -> age, discrete numeric
#A7_owner-A7_advanced -> skill/edu, discrete numeric
#A8_new -> income, discrete numeric
#A9 -> gender, discrete char
####################################################################
#Logistic regression to estimate probability of people plant hedgerows (shrubs, trees, grasses)

# caNew_2$new_UCCE <- factor(caNew_2$new_UCCE)
# caNew_2$new_RCD <- factor(caNew_2$new_RCD)
# caNew_2$new_NGO <- factor(caNew_2$new_NGO)
# caNew_2$new_CC <- factor(caNew_2$new_CC)
# caNew_2$new_Agencies <- factor(caNew_2$new_Agencies)
# caNew_2$X5q <- ordered(caNew_2$X5q)
# caNew_2$X6q <- ordered(caNew_2$X6q)
# caNew_2$Q7 <- ordered(caNew_2$Q7)
# caNew_2$A3 <- factor(caNew_2$A3)
# caNew_2$A4 <- factor(caNew_2$A4)
# caNew_2$A6 <- factor(caNew_2$A6)
# caNew_2$A7_owner <- factor(caNew_2$A7_owner)
# caNew_2$A7_technical <- factor(caNew_2$A7_technical)
# caNew_2$A7_bachelor <- factor(caNew_2$A7_bachelor)
# caNew_2$A7_advanced <- factor(caNew_2$A7_advanced)
# caNew_2$A8 <- factor(caNew_2$A8)
# caNew_2$A9 <- factor(caNew_2$A9)
# caNew_2$A2 <- as.numeric(caNew_2$A2)


str(caNew_2)
#Multiple logistic regression with demographic predictors
summary(lm.h <- glm(use1_h~benefit+concern+new_UCCE+new_RCD+new_NGO+new_CC+new_Agencies+X5q+X6q+Q7+A2+A3+A4+A6+A7_owner+A7_technical+A7_bachelor+A7_advanced+A8+A9,caNew_2,family=binomial))

#stepwise variable selection
library(MASS)
stepw <- stepAIC(lm.h, direction = "both")
summary(stepw)

#ordered x5q, x6q, Q7
caNew_2$X5q <- ordered(caNew_2$X5q)
caNew_2$X6q <- ordered(caNew_2$X6q)
caNew_2$Q7 <- ordered(caNew_2$Q7)

summary(lm.ho <- glm(use1_h~benefit+concern+new_UCCE+new_RCD+new_NGO+new_CC+new_Agencies+X5q+X6q+Q7+A2+A3+A4+A6+A7_owner+A7_technical+A7_bachelor+A7_advanced+A8+A9,caNew_2,family=binomial))

#stepwise variable selection
library(MASS)
stepw <- stepAIC(lm.h, direction = "both")
summary(stepw)

```


- Logistic regression to estimate probability of people plant wildflowers (forbs)

- Logistic regression to estimate probability of people plant native perennial grasses

2. Assumptions checking

